---
title: "hmwk6.2A"
author: "student"
date: "2023-09-10"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## CUSUM Part 1

Part 1: Using July through October daily-high-temperature data for Atlanta for 1996 through 2015, use
a CUSUM approach to identify when unofficial summer ends (i.e., when the weather starts
cooling off) each year.

**Summary**

*First things first, I eyeballed the temperature data and picked months that seemed to be pretty stable, year-over-year — July and August, in this case. I wanted to get a sense of what "normal" looks like, so I calculated the average high temperature for these two summer months for each year over a 20-year span. Think of this as my baseline or "critical value" (I'm calling it ref_value in the code).*

*Onto the hard part: setting the threshold! I played around with a bunch of numbers, but ultimately I used a range between -35 to -40, this is the CUSUM formula at play: cusum <- cusum + (yearly_data$Temperature[i] - ref_value). So it’s basically the running sum of how far off each day's temperature is from my average 'critical value.'*

*The kicker: when I looked at the results, it turned out that most years signaled the end of Atlanta's summer in September( maybe a clustering model at this point might be good!). This actually made sense because, when I checked out a time series plot of average high temperatures, there's a noticeable dip starting around August and going into September.*

*That's how I used CUSUM to decode Atlanta's end of summer days. The last pages is a running table of CUSUM calculations, just for because I was curious to see it in action.*

Load Libraries
```{r}
library(tidyr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(zoo)
```

## CUSUM
Get Data
```{r}
# construct the file path
file_path <- paste0("C:\\Users\\Public\\Documents\\gatech\\hw3-FA23\\data 6.2\\temps.txt")

# read in the data
weather_data <- read.table(file_path, header = TRUE)

#view data
tail(weather_data)
```

Aggregate Plot
```{r}
# convert DAY column to date format
weather_data$DAY <- as.Date(weather_data$DAY, format="%d-%b")

# convert to long format
long_data <- tidyr::pivot_longer(weather_data, cols=-DAY, names_to="Year", values_to="Temperature")

# extract month and add as new column
long_data$Month <- format(long_data$DAY, "%b")

# calculate monthly averages
monthly_avg <- dplyr::group_by(long_data, Year, Month)
monthly_avg <- dplyr::summarise(monthly_avg, Avg_Temperature = mean(Temperature, na.rm=TRUE))

# month as an ordered factor in monthly_avg
monthly_avg$Month <- factor(monthly_avg$Month, levels=c("Jul", "Aug", "Sep", "Oct"))

# plot monthly averages for each year
ggplot2::ggplot(monthly_avg, ggplot2::aes(x=Month, y=Avg_Temperature, group=Year, color=Year)) +
  ggplot2::geom_line() +
  ggplot2::labs(title="Monthly Average Temperatures", x="Month", y="Average Temperature (°F)")

# group by year and month, then calculate average temperature for each month
monthly_avg <- dplyr::group_by(long_data, Year, Month)
monthly_avg <- dplyr::summarise(monthly_avg, Avg_Temperature = mean(Temperature, na.rm=TRUE))

# calc average daily high and lowest average high temperature with month
min_max_avg_per_year <- dplyr::group_by(monthly_avg, Year)
min_max_avg_per_year <- dplyr::summarise(
  min_max_avg_per_year,
  Average_Daily_High = mean(Avg_Temperature, na.rm=TRUE),
  Min_AVG_Temperature = min(Avg_Temperature, na.rm=TRUE),
  Min_Month = Month[which.min(Avg_Temperature)],
  Max_AVG_Temperature = max(Avg_Temperature, na.rm=TRUE),
  Max_Month = Month[which.max(Avg_Temperature)]
)

# Print summary
print(paste("Metrics for each year:"))
print(min_max_avg_per_year)

```
Apply CUSUM
```{r}
# Initialize empty data frame to store end-of-summer days and CUSUM calculations
end_of_summer <- data.frame(Year=integer(), Day=as.Date(character()))
cusum_data <- data.frame(Year=integer(), Day=as.Date(character()), Temperature=numeric(), Ref_Value=numeric(), CUSUM=numeric())

# loop through each unique year in the dataset
for (year in unique(long_data$Year)) {
  
  # data for the current year
  yearly_data <- dplyr::filter(long_data, Year == year)
  
  # data for July and August for specific year
  jul_aug_data_year <- dplyr::filter(yearly_data, Month %in% c("July", "Aug"))
  
  # average high temperature for July and August for specific year
  ref_value <- mean(jul_aug_data_year$Temperature, na.rm=TRUE)
  
  # initialize CUSUM variable
  cusum <- 0
  
  # loop through each day in this year
  for (i in 1:nrow(yearly_data)) {
    cusum <- cusum + (yearly_data$Temperature[i] - ref_value)
    
    # add CUSUM data to cusum_data
    cusum_data <- rbind(cusum_data, data.frame(
      Year=year,
      Day=yearly_data$DAY[i],
      Temperature=yearly_data$Temperature[i],
      Ref_Value=ref_value,
      CUSUM=cusum
    ))
    
    # try different thresholds
    if (cusum < -40) {
      end_of_summer <- rbind(end_of_summer, data.frame(Year=year, Day=yearly_data$DAY[i]))
      break
    }
  }
}

# display the days that indicate the end of unofficial summer for each year
print("End of unofficial summer for each year:")
print(end_of_summer)

```

Plot the end of summer days
```{r}
# plot the end_of_summer data
ggplot(end_of_summer, aes(x=Year, y=Day)) +
  geom_point() +
  geom_line() +
  ggtitle("End of Unofficial Summer for Each Year") +
  xlab("Year") +
  ylab("Day")
```
View Full CUSUM Calculations
```{r}
# display the CUSUM calculations
print("CUSUM calculations:")
print(cusum_data)
```
